The truncated KdV framework for modeling anomalous waves induced by abrupt depth changes

### 2019-10-02 paper2_JNLS/paper.tex
NM

A recent series of laboratory \cite{bolles2019anomalous} and numerical investigations \cite{viotti2014} have demonstrated the emergence of anomalous wave statistics from abrupt variations in bottom topography. Since topographical variations are strictly one-dimensional, these studies can be viewed as offering a bare minimum set of conditions capable of generating anomalous wave statistics. In particular, the more complex mechanism of focusing by 2D bottom topography absent. The studies thus offer a paradigm system, with emergent anomalous statistical properties similar to that seen in more complex systems, but in a tractable context that is amenable to analysis.

Our particular focus is the laboratory experiments of Bolles et al.~\cite{bolles2019anomalous}, who demonstrated the emergence of anomalous statistics from a randomized wave-field encountering an abrupt depth change (ADC). In these experiments, the incoming wave-field is generated with nearly Gaussian statistics, and a plexiglass step placed near the middle of the tank creates the depth change. Upon passing into the shallower region, the wave distribution skews strongly towards positive displacement, with deviation from Gaussian being most pronounced a short distance downstream of the ADC. Inspired by these experimental observations, Majda et al.~\cite{majda2019statistical} developed a theoretical framework that accurately captures several key aspects of the anomalous behavior. The theory is based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) equations, and uses a combination of computational, statistical, and analytical tools. In follow-up work, Majda and Qi \cite{majdaqi2019} analyzed more severely truncated systems of lower dimension that were shown to exhibit many of the same anomalous properties while enjoying a more tractable structure.

The purpose of the present manuscript is to provide a more comprehensive treatment of the {\em combined} experimental and theoretical efforts of Bolles et al~\cite{bolles2019anomalous} and Majda et al.~\cite{majda2019statistical, majdaqi2019}. Special emphasis is placed on the synergy between experiment and theory -- a cooperative strategy with a proven record of success \cite{camassa2012stratified, ristroph2012, ganedi2018equilibrium}. We present an assortment of previously unpublished experimental data, for example on surface slope statistics and autocorrelation, which motivates further examination and refinement of the theory. Likewise, novel theoretical predictions have inspired us to perform additional experimental measurements, particularly regarding a relationship between surface-displacement skewness and surface-slope variance that emerges from the theory. Comparison with the new measurements further confirms the predictive power of the theoretical framework developed by Majda et al.~\cite{majda2019statistical}. In addition, we provide a more thorough treatment of the link between experimental and model parameters to facilitate comparison between the two. In particular, we flesh out details of non-dimensionalization, which were only briefly addressed in prior work for the sake of brevity.

@@

% Definitely cite several of Sapsis's new papers.

@@

	As discussed in \cite{bolles2019anomalous}, the experiments consist of a long, narrow wave tank (6 m long x 20 cm wide x 30 cm high), with waves generated by plexiglass paddle at one end. The waves propagate through the tank and, roughly midway through, pass over an abrupt depth change (ADC) created by a plexiglass step. The waves continue to propagate through the shallower depth until reaching the far end of the tank, at which point their energy is dissipated by a horse-hair dampener. Since the dampener minimizes the backscatter, the waves in this experiment propagate primarily in one direction, from left to right in Fig.~\ref{ExpDiagStats}.

The pivoting motion of the paddle is driven by a 5-phase stepper motor. To generate a randomized incoming wave field, the paddle angle $\phi$ is specified by a precomputed psuedo-random signal
\begin{align}
\label{PaddleAngle}
& \phi(t) = \phi_0 + \Delta \phi \sum_{n=1}^N a_n \cos(\omega_n t+\delta_n) \, , \\
\label{anEq}
& a_n = \sqrt{\frac{2 \Delta \omega}{\pi^{1/2} \omsig}} \, 
\exp \left( -\frac{(\omega_n - \omavg)^2}{2 \omsig^2} \right) \, ,
\end{align}
(Literal copy and paste) Here, the angular frequencies are evenly spaced $\omega_n = n  \Delta \omega$ with step size $ \Delta \omega = (\omavg+4 \omsig)/N$, where $\omavg$ and $\omsig$ represent the mean and the bandwidth of $\omega$ respectively. We set $\omavg = \omsig = 12.5$ rad/s, corresponding to a peak forcing frequency of 2 Hz and bandwidth of 2 Hz. Importantly, the phases $\delta_n$ are uniformly distributed random variables. We fix $N = 3000$, which sets a fundamental period of $T = 300$ seconds.

 % Figure: Experimental Diagram and Histograms
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.85 \linewidth]{Figs/ExpDiagStats.pdf}
\caption{
(a) Experimental schematic. Adapted from \cite{bolles2019anomalous}.
(b)--(c) Surface displacement measured at a representative upstream and downstream location. (d)--(e) Corresponding histograms.
}
%(b)--(c) Surface displacement measured at a representative upstream and downstream location. (d)--(e) Corresponding histograms.
\label{ExpDiagStats}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
 % Data from column 5 in MasterTimeSeries, Delta theta = 1.38 degrees, which is in between runs 5 and 6.

	The free surface is illuminated by light-emitting diodes running along the bottom of the tank and then imaged from the sideview with a Nikon D3300 at 60 frames per second. The illumination technique coupled with high pixel count of the camera allows surface displacements to be resolved with a accuracy better than 1/3 millimeter. Representative free surface measurements are shown in Figs.~\ref{ExpDiagStats}(b)--(c) at a position upstream and downstream of the ADC. The corresponding histograms are shown in Figs.~\ref{ExpDiagStats}(d)--(e).

TO DO: Describe the `representative' experiments in which $\dratdn = 0.24$ and $\epsup = 0.017$.

Show diagram, talk about different statistics on either side of ADC. Say Bolles found the downstream measurements to be remarkably well fit by a mean-zero gamma distribution.

@@

In this section, the variable-depth KdV equation \eqref{KdV} will be recast into a dimensionless form chosen for convenience in working with the statistical-mechanics framework introduced in \cite{majda2019statistical}. Since the choice of normalization not unique, it is instructive to first introduce a generic normalization to facilitate comparison with other choices. To this end, consider characteristic scales, $\ampscale, \lengthscale, \timescale$ for the wave amplitude, longitudinal length, and time respectively, which can remain unspecified for the moment. We introduce the dimensionless variables
\begin{align}
&u = \eta / \ampscale
&&\mbox{\em dimensionless surface displacement} \\
&\tilde{x} = \xi / \lengthscale
&&\mbox{\em dimensionless position (in moving frame)} \\
&\tilde{t} = t / \timescale
&&\mbox{\em dimensionless time}
\end{align}
Recasting \eqref{KdV} in terms of these variables gives the generic dimensionless KdV equation:
\begin{equation}
u_t + \frac{3}{2} \left( \frac{c \timescale \ampscale}{\lengthscale \depth} \right) u u_x 
+ \frac{1}{6} \left( \frac{c \timescale \depth^2}{\lengthscale^3} \right) u_{xxx} = 0
\end{equation}
We have dropped the tilde notation above for simplicity and will henceforth use tildes only in cases of possible ambiguity.

@@

A few comments are in order. First, we note that the original formulation of this theory utilized a slightly different normalization \cite{majda2019statistical}, with identical powers of $\drat$ in \eqref{dimlessKdV} but with different expressions for the other dimensionless parameters. These differences are purely cosmetic, and we have made the choices above simply to facilitate comparison against the experiments. Second, an alternate formulation of the variable-depth KdV equation exists in which it is conjectured that the product $\depth^{1/4} \eta$, rather than $\eta$, varies continuously across the ADC \cite{johnson1997modern}. Of course, that assumption implies a discontinuity in surface displacement, which, though small, is physically unrealistic. We have chosen to enforce continuity of surface displacement in the above on the basis of physical realism. Using the alternative formulation would simply modify the power of $\drat$ in the second term of \eqref{dimlessKdV} by 1/4, i.e.~$\drat^{-7/4}$ instead of $\drat^{-3/2}$. Hence, all of the results discussed below would still hold qualitatively in this alternate formulation, with only slight quantitative modifications.

%NOTE: Perhaps we should be more specific about comparing to the dimensionless parameters in our PNAS paper \cite{majda2019statistical}, BUT the expressions for E0 and L0 given in the PNAS paper do not match with what I calculate comparing to the above. Plus, it seems like it could only add confusion.

@@

The variable-depth KdV \eqref{dimlessKdV}, though not Hamiltonian throughout the entire domain, admits a Hamiltonian structure on each side of the ADC. Indeed, \eqref{dimlessKdV} can be expressed as
\begin{align}
\partial_t{u} = \sympJ \vard{\Hupdn}{u}
\end{align}
where $\sympJ = \pdi{}{x}$ is the symplectic operator and $\Ham = \Hupdn$ is the Hamiltonian, which will take different forms on either side of the ADC. It is convenient to decompose the Hamiltonian into a so-called cubic and quadratic component, given respectively by
\begin{align}
\label{H3H2}
\Hthree = \frac{1}{6} \int_{-\pi}^{\pi} u^3 \dx	\, , \qquad
\Htwo = \frac{1}{2} \int_{-\pi}^{\pi} u_x^2 \dx	\, .
\end{align}
Then the Hamiltonian can be expressed as
\begin{equation}
\label{Hamiltonian}
\Hupdn = C_2 \dratupdn^{1/2} \, \Htwo - C_3 \dratupdn^{-3/2} \, \Hthree
\end{equation}
where $\drat = \dratupdn$ changes value across the ADC. More explicitly, substituting \eqref{dratpw}, the separate upstream and downstream Hamiltonians are
\begin{align}
&\Hup = C_2 \, \Htwo - C_3 \, \Hthree 						&& \text{for } t<0 \\
&\Hdn = C_2 \dratdn^{1/2} \, \Htwo - C_3 \dratdn^{-3/2} \, \Hthree	&& \text{for } t>0
\end{align}
The above formulation differs cosmetically from some previous ones \cite{abramov2003hamiltonian, bajars2013weakly, majda2019statistical}, which used the opposite sign in both $\sympJ$ and $\Ham$. With that sign convention, Majda et al. 2019 found that a {\em negative} inverse temperature is required to accurately describe the experimental observations \cite{majda2019statistical}. We have chosen the sign convention above (which also happens to be consistent with Lax \cite{lax1975periodic}) so that a {\em positive} inverse temperature could be used and thus fit the theory into the most conventional form of statistical mechanics.

@@

Remarkably, the TKdV system \eqref{TKdV} retains nearly the same Hamiltonian structure described in Section \ref{HamiltonianSection}, with the only modification being the inclusion of the projection operator \cite{bajars2013weakly, majda2019statistical}. The piecewise defined Hamiltonian for TKdV is given by
\begin{equation}
\label{TruncHamiltonian}
\HLupdn = C_2 \dratupdn^{1/2} \, \Htwo[\uL] - C_3 \dratupdn^{-3/2} \, \Hthree[\uL]
\end{equation}
where $\Hthree$ and $\Htwo$ the defined exactly as before \eqref{H3H2}, but now are simply applied to the projected variable $\uL = \Proj u$.
Then TKdV \eqref{TKdV} can be expressed as
\begin{align}
\partial_t {\uL} = \partial_x \Proj \, \vard{\HLupdn}{\uL}
\end{align}
where the truncated symplectic operator is $\SympL = \partial_x \Proj$.
% Note: Bajars 2013 has some important details on the Hamiltonian on the truncated system.

@@

\vsp{5}
%----------------------------------------------------------------------------%
% Skewness formula
%----------------------------------------------------------------------------%
\subsection{Explicit formula for outgoing skewness}
We now examine an explicit formula for the outgoing skewness as originally founded in \cite{majda2019statistical}.
An experimental observation is that the incoming skewness is small, $\meanup{H_3} \approx 0$. Inserting this approximation into the statistical matching condition \eqref{statmatch} and simplifying gives
\begin{equation}
\label{H3H2}
\frac{\meandn{\Hthree^+}} {\meandn{\Htwo^+} - \meanup{\Htwo^+}} = \frac{C_2}{C_3} \dratdn^2
\end{equation}
%where $\Delta \mean{H_2} =  \meandn{H_2} - \meanup{H_2}$  indicates the difference between the upstream and downstream values.
To convert to dimensional, i.e.~experimental, values we first note that
\begin{align}
\mean{\Hthree^+} = \frac{\pi}{3} \mean{u^3} = 
\frac{1}{3} \pi^{-1/2} \frac{\mean{\eta^3}}{\etastd^3} = \frac{1}{3} \pi^{-1/2} \skw(\eta)
\end{align}
Second, to convert the downstream surface slope, we note
\begin{align}
\pd{u}{\tilde{x}} = \frac{\lamfac \lamdn}{\pi^{3/2} \etastd} \pd{\eta}{\xi}
\end{align}
where slope with respect to $\xi$ is the same as with respect to dimensional $x$.
Note: I believe it is important here to use $\lamdn$ in reverting to dimensional variables since we are matching $\Hdn$.
Hence, we have
\begin{align}
&\mean{\Htwo^+} = \pi \mean{ \left( \pd{u}{\tilde{x}} \right)^2} = 
\left( \frac{\lamfac \lamdn}{\pi \etastd} \right)^2 \var(\eta_x)
\end{align}
%
Then algebra gives
\begin{equation}
\frac{\meandn{\Hthree}} {\meandn{\Htwo} - \meanup{\Htwo}} = 
\frac{\pi^{3/2} \, \etastd^2} {3 \lamfac^2 \lamdn^2} 
\left( \frac{\skwdn(\eta)} { \vardn(\eta_x) - \varup(\eta_x)} \right)
\end{equation}
Also we have the simplification
\begin{equation}
\frac{C_2}{C_3} = \frac{\pi^{3/2} \delup^2}{9 \lamfac^2 \epsup}
\end{equation}
Hence, \eqref{H3H2} gives the remarkably explicit formula
\begin{equation}
\frac{\skwdn(\eta)} {\vardn(\eta_x) - \varup(\eta_x)} = \frac{1}{3} \epsup^{-3} \dratdn^3 =
\frac{1}{3} \left( \frac{\ddn}{\etastd} \right)^3
\end{equation}
In particular, the ratio on the left, which can be measured in the experiments, is predicted to scale as the inverse cube of the wave amplitude and as the square of the depth ratio.
% Note: previously I had D^2 on the RHS. The power of 3 comes from using lamdn consistently in the nondimensionalization.

>>>>

A recent series of laboratory \cite{} and numerical investigations \cite{viotti2014} have demonstrated the emergence of anomalous wave statistics from abrupt variations in bottom topography. Since topographical variations are strictly one-dimensional, these studies can be viewed as offering a bare minimum set of conditions capable of generating anomalous wave statistics. In particular, the more complex mechanism of focusing by 2D bottom topography absent. The studies thus offer a paradigm system, with emergent anomalous statistical properties similar to that seen in more complex systems, but in a tractable context that is amenable to analysis.

Our particular focus is the laboratory experiments of Bolles et al.~\cite{}, who demonstrated the emergence of anomalous statistics from a randomized wave-field encountering an abrupt depth change (ADC). In these experiments, the incoming wave-field is generated with nearly Gaussian statistics, and a plexiglass step placed near the middle of the tank creates the depth change. Upon passing into the shallower region, the wave distribution skews strongly towards positive displacement, with deviation from Gaussian being most pronounced a short distance downstream of the ADC. Inspired by these experimental observations, Majda et al.~\cite{majda2019} developed a theoretical framework that accurately captures several key aspects of the anomalous behavior. The theory is based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) equations, and uses a combination of computational, statistical, and analytical tools. In follow-up work, Majda and Qi \cite{majdaqi2019} analyzed more severely truncated systems of lower dimension that were shown to exhibit many of the same anomalous properties while enjoying a more tractable structure.

The purpose of the present manuscript is to provide a more comprehensive treatment of the {\em combined} experimental and theoretical efforts of Bolles et al~\cite{} and Majda et al.~\cite{majda2019, majdaqi2019}. Special emphasis is placed on the synergy between experiment and theory -- a cooperative strategy with a proven record of success \cite{camassa2012stratified, ristroph2012, ganedi2018equilibrium}. We present an assortment of previously unpublished experimental data, for example on surface slope statistics and autocorrelation, which motivates further examination and refinement of the theory. Likewise, novel theoretical predictions have inspired us to perform additional experimental measurements, particularly regarding a peculiar relationship between surface-displacement skewness and surface-slope variance that is predicted by the theory. Comparison with the new measurements further confirms the predictive power of the framework developed by Majda et al.~\cite{majda2019}. In addition, we provide a more thorough treatment of the link between experimental and model parameters to facilitate comparison between the two. In particular, we flesh out details of non-dimensionalization, which were only briefly addressed in prior work for the sake of brevity.

@@

@@

	As diagrammed in Fig.~\ref{ExpDiagStats}(a), the experiments consist of a long, narrow wave tank (6 m long x 20 cm wide x 30 cm high), with waves generated by plexiglass paddle at one end \cite{bolles2019}. The waves propagate through the tank and, roughly midway through, pass over an abrupt depth change (ADC) created by a plexiglass step. The waves continue to propagate through the shallower depth until reaching the far end of the tank, at which point their energy is dissipated by a horse-hair dampener. Since the dampener minimizes backscatter, the waves in this experiment propagate primarily in one direction, from left to right in Fig.~\ref{ExpDiagStats}(a).

The pivoting motion of the paddle is driven by a 5-phase stepper motor. To generate a randomized incoming wave field, the paddle angle $\phi$ is specified by a precomputed psuedo-random signal
\begin{align}
\label{PaddleAngle}
& \phi(t) = \phi_0 + \Delta \phi \sum_{n=1}^N a_n \cos(\omega_n t+\delta_n) \, , \\
\label{anEq}
& a_n = \sqrt{\frac{2 \Delta \omega}{\pi^{1/2} \omsig}} \, 
\exp \left( -\frac{(\omega_n - \omavg)^2}{2 \omsig^2} \right) \, ,
\end{align}
The angular frequencies are evenly spaced $\omega_n = n  \Delta \omega$ with step size $ \Delta \omega = (\omavg+4 \omsig)/N$, where $\omavg$ and $\omsig$ represent the mean and the bandwidth of $\omega$ respectively. As in prior work, all experiments reported here use the values $\omavg = \omsig = 12.5$ rad/s, corresponding to a peak forcing frequency of 2 Hz and bandwidth of 2 Hz. Meanwhile, the phases $\delta_n$ are uniformly distributed random variables.

% Figure: Experimental Diagram and Histograms
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.85 \linewidth]{Figs/ExpDiagStats.pdf}
\caption{
(a) Experimental schematic. 
(b)--(c) Surface displacement measured at a representative upstream and downstream location. (d)--(e) Corresponding histograms. Figure adapted from \cite{bolles2019}.}
\label{ExpDiagStats}
\end{center}
\end{figure}
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Data from column 5 in MasterTimeSeries, Delta theta = 1.38 degrees, which is in between runs 5 and 6.

	The free surface is illuminated by light-emitting diodes that run along the bottom of the tank and is then imaged from the sideview with a Nikon D3300 at 60 frames per second. The illumination technique, coupled with high pixel count of the camera, allows surface displacements to be resolved with accuracy better than 1/3 millimeter. Furthermore, these optical measurements permit extraction of wave statistics {\em continuously} in space, rather than at a few discrete locations. Further details of the experimental setup can be found in \cite{bolles2019}.

	Example measurements of free-surface displacements $\eta$ are shown in Figs.~\ref{ExpDiagStats}(b)--(c). These measurements are taken at two representative locations: one a short distance upstream of the ADC and one downstream of the ADC. Both signals exhibit a combination of periodic and random fluctuations, with the dominant oscillations corresponding to the peak driving frequency of 2 Hz. The nature of the random fluctuations is revealed by the corresponding histograms shown in Fig.~\ref{ExpDiagStats}(d)--(e). The upstream measurements are symmetrically distributed about the mean displacement of $\eta = 0$. In fact, Bolles et al.~found that these measurements closely follow a Gaussian distribution \cite{bolles2019}. The downstream measurements, however, are not symmetrically distributed and skew strongly towards positive displacement, $\eta > 0$. Bolles et al.~found that these measurements can be well approximated by a mean-zero gamma distribution \cite{bolles2019}.

\vsp{5}
TO DO: Describe the `representative' experiments in which $\dratdn = 0.24$ and $\epsup = 0.017$.

 % More details in \cite{}

@@

In this section, the variable-depth KdV equation \eqref{KdV} will be recast into a dimensionless form chosen for convenience in working with the statistical-mechanics framework introduced in \cite{majda2019}. Since the choice of normalization not unique, it is instructive to first introduce a generic normalization to facilitate comparison with other choices. To this end, consider characteristic scales, $\ampscale, \lengthscale, \timescale$ for the wave amplitude, longitudinal length, and time respectively, which can remain unspecified for the moment. We introduce the dimensionless variables
\begin{align}
&u = \eta / \ampscale
&&\mbox{\em dimensionless surface displacement} \\
&\tilde{x} = \xi / \lengthscale
&&\mbox{\em dimensionless position (in moving frame)} \\
&\tilde{t} = t / \timescale
&&\mbox{\em dimensionless time}
\end{align}
Recasting \eqref{KdV} in terms of these variables gives the generic dimensionless KdV equation:
\begin{equation}
u_t + \frac{3}{2} \left( \frac{c \timescale \ampscale}{\lengthscale \depth} \right) u u_x 
+ \frac{1}{6} \left( \frac{c \timescale \depth^2}{\lengthscale^3} \right) u_{xxx} = 0
\end{equation}
We have dropped the tilde notation above for simplicity and will henceforth use tildes only in cases of possible ambiguity.

@@

A few comments are in order. First, we note that the original formulation of this theory utilized a slightly different normalization \cite{majda2019}, with identical powers of $\drat$ in \eqref{dimlessKdV} but with different expressions for the other dimensionless parameters. These differences are purely cosmetic, and we have made the choices above simply to facilitate comparison against the experiments. Second, an alternate formulation of the variable-depth KdV equation exists in which it is conjectured that the product $\depth^{1/4} \eta$, rather than $\eta$, varies continuously across the ADC \cite{johnson1997modern}. Of course, that assumption implies a discontinuity in surface displacement, which, though small, is physically unrealistic. We have chosen to enforce continuity of surface displacement in the above on the basis of physical realism. Using the alternative formulation would simply modify the power of $\drat$ in the second term of \eqref{dimlessKdV} by 1/4, i.e.~$\drat^{-7/4}$ instead of $\drat^{-3/2}$. Hence, all of the results discussed below would still hold qualitatively in this alternate formulation, with only slight quantitative modifications.

%NOTE: Perhaps we should be more specific about comparing to the dimensionless parameters in our PNAS paper \cite{majda2019}, BUT the expressions for E0 and L0 given in the PNAS paper do not match with what I calculate comparing to the above. Plus, it seems like it could only add confusion.

@@

The variable-depth KdV \eqref{dimlessKdV}, though not Hamiltonian throughout the entire domain, admits a Hamiltonian structure on each side of the ADC. Indeed, \eqref{dimlessKdV} can be expressed as
\begin{align}
\partial_t{u} = \sympJ \vard{\Hupdn}{u}
\end{align}
where $\sympJ = \pdi{}{x}$ is the symplectic operator and $\Ham = \Hupdn$ is the Hamiltonian, which will take different forms on either side of the ADC. It is convenient to decompose the Hamiltonian into a so-called cubic and quadratic component, given respectively by
\begin{align}
\label{H3H2}
\Hthree = \frac{1}{6} \int_{-\pi}^{\pi} u^3 \dx	\, , \qquad
\Htwo = \frac{1}{2} \int_{-\pi}^{\pi} u_x^2 \dx	\, .
\end{align}
Then the Hamiltonian can be expressed as
\begin{equation}
\label{Hamiltonian}
\Hupdn = C_2 \dratupdn^{1/2} \, \Htwo - C_3 \dratupdn^{-3/2} \, \Hthree
\end{equation}
where $\drat = \dratupdn$ changes value across the ADC. More explicitly, substituting \eqref{dratpw}, the separate upstream and downstream Hamiltonians are
\begin{align}
&\Hup = C_2 \, \Htwo - C_3 \, \Hthree 						&& \text{for } t<0 \\
&\Hdn = C_2 \dratdn^{1/2} \, \Htwo - C_3 \dratdn^{-3/2} \, \Hthree	&& \text{for } t>0
\end{align}
The above formulation differs cosmetically from some previous ones \cite{abramov2003hamiltonian, bajars2013weakly, majda2019}, which used the opposite sign in both $\sympJ$ and $\Ham$. With that sign convention, Majda et al. 2019 found that a {\em negative} inverse temperature is required to accurately describe the experimental observations \cite{majda2019}. We have chosen the sign convention above (which also happens to be consistent with Lax \cite{lax1975periodic}) so that a {\em positive} inverse temperature could be used and thus fit the theory into the most conventional form of statistical mechanics.

@@

Remarkably, the TKdV system \eqref{TKdV} retains nearly the same Hamiltonian structure described in Section \ref{HamiltonianSection}, with the only modification being the inclusion of the projection operator \cite{bajars2013weakly, majda2019}. The piecewise defined Hamiltonian for TKdV is given by
\begin{equation}
\label{TruncHamiltonian}
\HLupdn = C_2 \dratupdn^{1/2} \, \Htwo[\uL] - C_3 \dratupdn^{-3/2} \, \Hthree[\uL]
\end{equation}
where $\Hthree$ and $\Htwo$ the defined exactly as before \eqref{H3H2}, but now are simply applied to the projected variable $\uL = \Proj u$.
Then TKdV \eqref{TKdV} can be expressed as
\begin{align}
\partial_t {\uL} = \partial_x \Proj \, \vard{\HLupdn}{\uL}
\end{align}
where the truncated symplectic operator is $\SympL = \partial_x \Proj$.
% Note: Bajars 2013 has some important details on the Hamiltonian on the truncated system.

@@

\vsp{5}
%----------------------------------------------------------------------------%
% Skewness formula
%----------------------------------------------------------------------------%
\subsection{Explicit formula for outgoing skewness}
We now examine an explicit formula for the outgoing skewness as originally founded in \cite{majda2019}.
An experimental observation is that the incoming skewness is small, $\meanup{H_3} \approx 0$. Inserting this approximation into the statistical matching condition \eqref{statmatch} and simplifying gives
\begin{equation}
\label{H3H2}
\frac{\meandn{\Hthree^+}} {\meandn{\Htwo^+} - \meanup{\Htwo^+}} = \frac{C_2}{C_3} \dratdn^2
\end{equation}
%where $\Delta \mean{H_2} =  \meandn{H_2} - \meanup{H_2}$  indicates the difference between the upstream and downstream values.
To convert to dimensional, i.e.~experimental, values we first note that
\begin{align}
\mean{\Hthree^+} = \frac{\pi}{3} \mean{u^3} = 
\frac{1}{3} \pi^{-1/2} \frac{\mean{\eta^3}}{\etastd^3} = \frac{1}{3} \pi^{-1/2} \skw(\eta)
\end{align}
Second, to convert the downstream surface slope, we note
\begin{align}
\pd{u}{\tilde{x}} = \frac{\lamfac \lamdn}{\pi^{3/2} \etastd} \pd{\eta}{\xi}
\end{align}
where slope with respect to $\xi$ is the same as with respect to dimensional $x$.
Note: I believe it is important here to use $\lamdn$ in reverting to dimensional variables since we are matching $\Hdn$.
Hence, we have
\begin{align}
&\mean{\Htwo^+} = \pi \mean{ \left( \pd{u}{\tilde{x}} \right)^2} = 
\left( \frac{\lamfac \lamdn}{\pi \etastd} \right)^2 \var(\eta_x)
\end{align}
%
Then algebra gives
\begin{equation}
\frac{\meandn{\Hthree}} {\meandn{\Htwo} - \meanup{\Htwo}} = 
\frac{\pi^{3/2} \, \etastd^2} {3 \lamfac^2 \lamdn^2} 
\left( \frac{\skwdn(\eta)} { \vardn(\eta_x) - \varup(\eta_x)} \right)
\end{equation}
Also we have the simplification
\begin{equation}
\frac{C_2}{C_3} = \frac{\pi^{3/2} \delup^2}{9 \lamfac^2 \epsup}
\end{equation}
Hence, \eqref{H3H2} gives the remarkably explicit formula
\begin{equation}
\frac{\skwdn(\eta)} {\vardn(\eta_x) - \varup(\eta_x)} = \frac{1}{3} \epsup^{-3} \dratdn^3 =
\frac{1}{3} \left( \frac{\ddn}{\etastd} \right)^3
\end{equation}
In particular, the ratio on the left, which can be measured in the experiments, is predicted to scale as the inverse cube of the wave amplitude and as the square of the depth ratio.
% Note: previously I had D^2 on the RHS. The power of 3 comes from using lamdn consistently in the nondimensionalization.
